SPANG: A SPARQL Client with One-Liners, Template Libraries, and Programming Interfaces

### 2020-09-29 paper.tex
add copyright
fix citation

@@

\title{SPANG: A SPARQL Client with One-Liners, Template Libraries, and Programming Interfaces}
\titlerunning{SPANG: A SPARQL Client with One-Liners, Template Libraries}
\author{Hirokazu Chiba\orcidID{0000-0003-4062-8903}}
\authorrunning{H. Chiba}
\institute{Database Center for Life Science, Chiba 277-0871, Japan\\
\email{chiba@dbcls.rois.ac.jp}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 15--250 words.
%How can we maximize the value of accumulated SPARQL queries? 
SPARQL is a key component of the Semantic Web, and thus, the reusability of SPARQL is crucial for maximizing productivity on the Semantic Web.
While an increasing number of datasets have been published in the Resource Description Framework (RDF) and the SPARQL standard allows for an interoperable framework for querying the RDF datasets, the reuse of SPARQL queries remains limited. We present SPANG, a framework to make SPARQL queries more reusable using a templating engine, one-liner functionality, and an application programming interface. Here we demonstrate its use cases for making SPARQL queries more reusable; assigning a set of metadata for annotating a SPARQL query, which enables to identify SPARQL query and provides useful information for the query; parameterizing SPARQL queries, which makes each query reusable under different settings; assigning a globally unique and stable identifier for each query; and constructing a server to provide indices of SPARQL queries. SPANG makes SPARQL queries more reusable in the local system or across the Web and leads to the construction of an eco-system wherein SPARQL queries are shared to maximize the productivity of Semantic Web developers and users. SPANG is available at \url{https://spang.dbcls.jp/}.

\keywords{SPARQL \and template engine \and one-liner \and query library}

@@

We implemented the SPANG template engine in JavaScript, where the template is transformed into an abstract syntax tree, processed using the given parameters, and then reconstructed into SPARQL to be submitted to an endpoint. 
We used Parsing Expression Grammar (PEG)~\cite{peg} expression of the SPARQL grammar based on the EBNF of the SPARQL specifications~\cite{sparql}. The original implementation of PEG.js~\cite{pegjs} of the SPARQL grammar exists in the rdfstore-js project (\url{https://github.com/antoniogarrote/rdfstore-js})~\cite{rdfstore-js}.
The PEG.js code was modified to fit the use in the case of the SPANG framework.
PEG.js is a parser generator, and thus, we could generate a custom parser based on the PEG.js description. 
%Thus, we obtained abstract syntax tree in the form of JSON object.
While other efforts have been made to implement the parser and formatter of SPARQL~\cite{sparql-js}, we developed a parser and formatter that stringify the generated abstract syntax tree based on the PEG expression. %Thus, SPANG can also reformat the input SPARQL.

@@

\bibitem{rdfstore-js}
Hernández, A. G., & GARCıA, M. N. M. (2012). A JavaScript RDF store and application library for linked data client applications. In Devtracks of the, WWW2012, conference. Lyon, France.

>>>>

\newcommand\blfootnote[1]{%
  \begingroup
  \renewcommand\thefootnote{}\footnote{#1}%
  \addtocounter{footnote}{-1}%
  \endgroup
}

@@

\title{SPANG: A SPARQL Client with One-Liners, Template Libraries, and Programming Interfaces}
\titlerunning{SPANG: A SPARQL Client with One-Liners, Template Libraries}
\author{Hirokazu Chiba\orcidID{0000-0003-4062-8903}}
\authorrunning{H. Chiba}
\institute{Database Center for Life Science, Chiba 277-0871, Japan\\
\email{chiba@dbcls.rois.ac.jp}
}
%
\maketitle              % typeset the header of the contribution
%

\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 15--250 words.
%How can we maximize the value of accumulated SPARQL queries? 
SPARQL is a key component of the Semantic Web, and thus, the reusability of SPARQL is crucial for maximizing productivity on the Semantic Web.
While an increasing number of datasets have been published in the Resource Description Framework (RDF) and the SPARQL standard allows for an interoperable framework for querying the RDF datasets, the reuse of SPARQL queries remains limited. We present SPANG, a framework to make SPARQL queries more reusable using a templating engine, one-liner functionality, and an application programming interface. Here we demonstrate its use cases for making SPARQL queries more reusable; assigning a set of metadata for annotating a SPARQL query, which enables to identify SPARQL query and provides useful information for the query; parameterizing SPARQL queries, which makes each query reusable under different settings; assigning a globally unique and stable identifier for each query; and constructing a server to provide indices of SPARQL queries. SPANG makes SPARQL queries more reusable in the local system or across the Web and leads to the construction of an eco-system wherein SPARQL queries are shared to maximize the productivity of Semantic Web developers and users. SPANG is available at \url{https://spang.dbcls.jp/}.
\blfootnote{Copyright © 2020 for this paper by its author. Use permitted under Creative Commons License Attribution 4.0 International (CC BY 4.0).}
\keywords{SPARQL \and template engine \and one-liner \and query library}

@@

We implemented the SPANG template engine in JavaScript, where the template is transformed into an abstract syntax tree, processed using the given parameters, and then reconstructed into SPARQL to be submitted to an endpoint. 
We used Parsing Expression Grammar (PEG)~\cite{peg} expression of the SPARQL grammar based on the EBNF of the SPARQL specifications~\cite{sparql}. The original implementation of PEG.js~\cite{pegjs} of the SPARQL grammar exists in the rdfstore-js project (\url{https://github.com/antoniogarrote/rdfstore-js}).
The PEG.js code was modified to fit the use in the case of the SPANG framework.
PEG.js is a parser generator, and thus, we could generate a custom parser based on the PEG.js description. 
%Thus, we obtained abstract syntax tree in the form of JSON object.
While other efforts have been made to implement the parser and formatter of SPARQL~\cite{sparql-js}, we developed a parser and formatter that stringify the generated abstract syntax tree based on the PEG expression. %Thus, SPANG can also reformat the input SPARQL.

@@
